% ============================================================
%  Appendix A — Configuration Quick Reference
% ============================================================
\chapter{Configuration Quick Reference} \label{app:config}

This appendix provides a compact one-page summary of every configurable
option in the template.  Use it as a cheat-sheet once you have read
Chapter~\ref{chap:template}.

% ------------------------------------------------------------
\section{Cover Metadata (\texttt{config/cover.tex})} \label{app:cover-ref}
% ------------------------------------------------------------

\begin{description}
  \item[\texttt{\textbackslash coverlanguage}]
        \texttt{en} or \texttt{pt}.  Controls cover language and selects
        the matching cover PDF asset.

  \item[\texttt{\textbackslash thesistitle}]
        Full thesis title.

  \item[\texttt{\textbackslash thesisauthor}]
        Author full name.

  \item[\texttt{\textbackslash thesisdegree}]
        Degree being awarded.

  \item[\texttt{\textbackslash thesissubmissiondate}]
        Month and year (e.g.\ \texttt{March 2026}).

  \item[\texttt{\textbackslash thesisschoolEN}]
        English school name.

  \item[\texttt{\textbackslash thesisschoolPT}]
        Portuguese school name.

  \item[\texttt{\textbackslash thesisuniversity}]
        University name.

  \item[\texttt{\textbackslash thesissupervisors}]
        Primary supervisor block.  Use \texttt{\textbackslash\textbackslash}
        for line breaks and \texttt{\textbackslash textit} for affiliation.

  \item[\texttt{\textbackslash thesiscosupervisors}]
        Co-supervisor block, same format.  Leave body empty (\texttt{\{\}})
        if not applicable.

  \item[\texttt{\textbackslash printbackcovertrue}]
        Enables the back cover.  Replace with
        \texttt{\textbackslash printbackcoverfalse} for digital-only.
\end{description}

% ------------------------------------------------------------
\section{Document Structure (\texttt{config/files.tex})} \label{app:files-ref}
% ------------------------------------------------------------

\begin{description}
  \item[\texttt{\textbackslash perchapterbibliographytrue/false}]
        Per-chapter vs.\ single end-of-document bibliography.

  \item[\texttt{\textbackslash printlistoffigurestrue/false}]
        Enables / disables the List of Figures.

  \item[\texttt{\textbackslash printlistoftablestrue/false}]
        Enables / disables the List of Tables.

  \item[\texttt{\textbackslash printappendicestrue/false}]
        Enables / disables the appendix block.

  \item[\texttt{\textbackslash printannexestrue/false}]
        Enables / disables the annex block.

  \item[\texttt{\textbackslash loadchapters}]
        Macro listing all main chapters via
        \texttt{\textbackslash chapterwithrefs\{path\}}.

  \item[\texttt{\textbackslash loadappendices}]
        Macro listing appendix files.  Must call
        \texttt{\textbackslash imsstartappendices} once at the top.

  \item[\texttt{\textbackslash loadannexes}]
        Macro listing annex files.  Must call
        \texttt{\textbackslash imsstartannexes} once at the top.
\end{description}

% ------------------------------------------------------------
\section{Front Matter (\texttt{chapters/preface.tex})} \label{app:preface-ref}
% ------------------------------------------------------------

\begin{description}
  \item[\texttt{\textbackslash setsdglanguage\{en|pt\}}]
        Language for SDG icon labels.

  \item[\texttt{\textbackslash setsdgs\{1,3,11\}}]
        Comma-separated list of SDG goal numbers to display (1--17).

  \item[\texttt{\textbackslash prefacesection\{Name\}}]
        Opens a new front-matter section.  Available section names:
        Statement of Integrity, Dedication, Acknowledgements, Quote,
        Abstract, Resumo, Glossary, Acronyms, Symbols.

  \item[\texttt{\textbackslash keywords\{kw1, kw2\}}]
        Used inside the Abstract section to print the keyword line.

  \item[\texttt{\textbackslash printsustainablegoals}]
        Inserts the SDG icon grid.  Call this at the end of the Abstract
        section.
\end{description}

% ------------------------------------------------------------
\section{Suggested Git Workflow} \label{app:workflow}
% ------------------------------------------------------------

Version control is strongly recommended.  A minimal workflow:

\begin{enumerate}
  \item Initialise: \texttt{git init} (or clone this repository).
  \item Work on a feature branch: \texttt{git checkout -b chapter3}.
  \item Commit frequently with descriptive messages:
\begin{verbatim}
git add chapters/chapter3.tex figures/ch3-*.pdf
git commit -m "chapter3: add methodology section"
\end{verbatim}
  \item Merge when stable: \texttt{git checkout main \&\& git merge chapter3}.
  \item Never commit generated files (\texttt{*.aux}, \texttt{*.bbl},
        \texttt{main.pdf}, etc.).  The \texttt{.gitignore} already excludes
        the most common ones.
\end{enumerate}

Before submission, tag the final version:

\begin{verbatim}
git tag -a v1.0 -m "Final submission"
git push origin v1.0
\end{verbatim}

